---
title: "Graffiti mitigation"
author: 
- |
    | Student number: TEST
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output: html_document
---
# Originality declaration  

I, [**Xiaoyi Chen**], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

date: `r format(Sys.time(), '%d %B, %Y')`

# Start your response here

## Initial project scope
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(tmap)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(sf)
library(sp)
library(spdep)
library(car)
library(fs)
library(janitor)
library(here)
```


```{r}
#shape <- st_read("Data_graffiti/Census 2020_ Tracts for San Francisco/geo_export_b5c6b5d6-853b-4740-aa7d-7bc83ebe8e41.shp") %>%
#  st_transform(., crs=7131)

shape <- st_read("Data_graffiti/Census 2010 Tracts/geo_export_eaefa3d5-61aa-4b6a-876d-983ff8bd7b3f.shp") %>%
  st_transform(., crs=7131)

graffiti <- read_csv("Data_graffiti/Graffiti.csv")

graffiti2<-graffiti%>%
  separate(., Point, c("latitude", "longitude"), sep = ",")
# parse makes a string containing a number into a number  
graffiti2$latitude<-parse_number(graffiti2$latitude) ## leading $ and grouping character , ignored
graffiti2$longitude<-parse_number(graffiti2$longitude) ## leading $ and grouping character , ignored

graffiti3<- graffiti2%>%
  # if i don't remove NAs or 0 it may cause an error - remember last week
  filter(latitude !=	0 )%>%
  filter(longitude != 0)%>%
  st_as_sf(., coords = c("longitude", "latitude"), 
                   crs = 4326)

graffiti4<- graffiti3%>%
filter(str_detect(Closed, "2019"))%>%
  #filter(str_detect(`Request Type`, "Building"))%>%
  st_transform(., crs=7131)

# spatial subset points within the san fran polygon (sf object)

graffiti_within <- graffiti4[shape, ,op=st_intersects]

tmap_mode("plot")
tm_shape(shape) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(graffiti4) +
  tm_dots(col = "blue")

```



```{r, eval=FALSE}

# 

  points_sf_joined <- shape%>%
    st_join(graffiti4)%>%
    add_count(geoid10)%>%
    janitor::clean_names()%>%
    #calculate area
    mutate(area=st_area(.))%>%
    #then density of the points per ward
    mutate(density=n/area)%>%
    dplyr::select(geoid10 , neighborhood, density)%>%
    group_by(geoid10) %>%         
  summarise(geoid10 = first(geoid10),
          neighborhood= first(neighborhood),
          density= first(density))


```




```{r}
################### joining examples
# the word predicate in the output means to meet a relationship (e.g. intersecting)
intersectexample <- st_intersects(shape, graffiti4)

# this is very similar to spatial join st_join(shape, graffiti4, op=st_intersects)
intersectionexample <- st_intersection(shape, graffiti4)

# why are there 5 extra points here? because 5 points don't intersect the shape!!
joinexample<-st_join(graffiti4, shape, join=st_intersects)

tmap_mode("plot")
tm_shape(intersectionexample) +
  tm_dots(col = NA, alpha = 0.5)
#####################
```

```{r}
points_sf_joined<-shape%>%
  mutate(n = lengths(st_intersects(., graffiti4)))%>%
  janitor::clean_names()%>%
    #calculate area
    mutate(area=st_area(.))%>%
    #then density of the points per ward
    mutate(density=n/area)

```



```{r}
#census_health <- read_csv("Data_graffiti/ACSST5Y2020.S2701_2022-12-06T131819/ACSST5Y2020.S2701-Data.csv", skip=1)
census_health <- read_csv("Data_graffiti/ACSST5Y2021.S2703_2023-12-13T175342/ACSST5Y2021.S2703-Data.csv", skip=1)
  
 census_health2 <- census_health%>%
  clean_names()
 
```


Within this data there is a column called "estimate_total_civilian_noninstitutionalized_population_coverage_alone_or_in_combination_employer_based_health_insurance_alone_or_in_combination_under_19", so here the study object is juvenile population who has health insurance.

Select the column then join the data to our main sf file...remember i am trying to model the density of graffiti here...

Note i am selecting characters from the geography column based on their position. More information on this can be found on the census website. But essentially the US stands for United States then next 2 are the state, the next 3 the county, we need the census tract version here: STATE (2 digits )+COUNTY (3 digits) +TRACT(6 digits)

```{r}

census_health3 <-census_health2 %>%
  select(geography, geographic_area_name,
         estimate_total_civilian_noninstitutionalized_population_coverage_alone_or_in_combination_employer_based_health_insurance_alone_or_in_combination_under_19)%>%
  # 10 is where the geoid starts
  mutate(tract = str_sub(geography, start = 10, end=20))



points_sf_joined_join_health_insur <- points_sf_joined %>%
  left_join(., census_health3,
            by=c("geoid10" = "tract"))

```

Or we can use tidy census to read in the census data directly - no joining.

Note, S2701_C03_031E is the estimate_percent_insured_civilian_noninstitutionalized_population, however it doesn't appear when i search for the data, this could be to differences in the API and the data online.

Instead i could extract the population per tract (B01003_001) and get a count of insured people (B27001_001) or i can just set the variable argument to S2701_C03_031E

```{r}
library(tidycensus)
options(tigris_use_cache = TRUE)

v20 <- load_variables(2020, "acs5")

View(v20)

census_SF <- get_acs(
  state = "CA",
  county = "San Francisco",
  geography = "tract",
  variables = c("B27001_001", "B27001_001"),
  geometry = TRUE,
  year = 2020
)%>%
    st_transform(., crs=7131)

head(census_SF)

```

No we need to clip this to our study area, at the moment it is for all the tracts in California.

Options:

1.  Drop geometry of the census data (from tidycensus) and left join to our tracts
2.  Use a spatial join (e.g. below)
3.  Work out what to filter on or adjust code above to include county - this was is the line county= San Francisco

```{r}
join <- st_join(shape, census_SF, join=st_equals)
```



```{r}
census_income <- read_csv("Data_graffiti/ACSST5Y2021.S1901_2023-12-14T065813/ACSST5Y2021.S1901-Data.csv", skip=1)
# ACSST5Y2020.S1901_2022-12-07T041746/ACSST5Y2020.S1901-Data.csv
  
census_income2 <- census_income%>%
  clean_names()
 

census_income3 <-census_income2 %>%
  select(geography, geographic_area_name, estimate_households_mean_income_dollars, estimate_households_median_income_dollars, estimate_households_total)%>%
  mutate(tract = str_sub(geography, start = 10, end=20))
 
 
```



```{r}
points_sf_joined_join_health_insur_income <- points_sf_joined_join_health_insur %>%
  left_join(., census_income3,
            by=c("geoid10" = "tract"))

```



